Modelling the Retinotectal projection - can competition provide a stopping mechanism?

### 2021-09-10 paper/paper.tex
WIP, switch to laptop

% Allow includegraphics and nice wrapped figures
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage[outercaption]{sidecap}

@@

At each timestep in a simulation, the position, $\mathbf{x}_{b,t}$, at time
$t$ of each branch $b$ is updated according to
%
\begin{equation}
\mathbf{x}_{b,t+1} = \mathbf{x}_{b,t} + \mathbf{M}_{b}
\end{equation}
%
where $\mathbf{M}_{b}$, the movement vector, is the sum of a chemoaffinity
effect, $\mathbf{G}_b$, axon-axon competition, $\mathbf{C}_b$, an axon-axon
receptor-receptor interaction, $\mathbf{I}_b$, and axon-axon receptor-ligand
interaction, $\mathbf{J}_b$, and a `border effect',
$\mathbf{B}_b$, which acts to retain branches within the tectal region:
%
\begin{equation}
\mathbf{M}_{b} = m_1 \mathbf{G}_b + m_2 \mathbf{C}_b + m_3 \mathbf{I}_b +
m_4 \mathbf{J}_b + m_5 \mathbf{B}_b
\end{equation}
%
evaluated at time $t$. $m_1$--$m_4$ are scalar parameters. $m_2$

@@

\subsection*{Axon-axon competition}

Competition, in the context of this model, is a repulsive effect which acts
between any two branches, whether they originate from different retinal
ganglion cells or from the same cell. It assumes that there is some kind of
self-other signal (\textbf{be sure to discuss possible origins}) which acts to
discourage two cells from approaching each other.
%
\begin{equation}
\mathbf{C}_b = \frac{1}{|B_b|} \sum_k \hat{\mathbf{x}}_{kb} W
\end{equation}
where the distance based weight $W$ is given by
\begin{equation}
W(d_{kb}) = \begin{cases}
      1 - \frac{d_{kb}}{2r}   & d_{kb} \leq 2r \\
     0 & d_{kb} > 2r
     \end{cases}
\end{equation}
for two growth cones of radius $r$ separated by a distance $d_{kb}$. $B_b$ is
the set of growth cones within a distance $2r$ of branch $b$.

@@

For axon-axon interactions, repulsion causes a movement of branch $b$ along a
unit vector, $\hat{\mathbf{x}}_{kb}$, from branch $k$ to branch $b$ causing
them to move further apart; attraction causes the opposite movement. The
interaction, $\mathbf{I}_b$, acting on branch $b$, is given by
%
\begin{equation}
% No \frac{1}{|B_b|} in this one
\mathbf{I}_b = \sum_k \hat{\mathbf{x}}_{kb}\,Q(d_{kb})
\end{equation}
%
where $Q(d_{kb})$ is the signalling strength between two growth cones of
radius $r$ a distance $d_{kb}$ from one another, given by
%
\begin{equation}
Q(d_{kb}) = \begin{cases}
     \sum_i^N F_i\,r_{i,b}\,l_{i,k}    & d_{kb} \leq 2r \\
     0 & d_{kb} > 2r
     \end{cases}
\end{equation}
%
The sign of $Q$ (which is dependent on the values of $F_i$) determines whether
the interaction, $\mathbf{I}_b$, is repulsive or attractive. $l_{i,k}$ is the
expression of ligand $i$ on branch $k$.

@@

\begin{figure}
\includegraphics[width=\linewidth]{./images/j4_ee_wt_fig.png}
\caption{Gradient following model with axon-axon interactions.}
\label{f:wt}
\end{figure}

@@

This result shows that a gradient following model based on receptor-ligand
interactions and incorporating axon-axon repulsion is able to generate the
basic mapping observed in the retinotectal projection.

@@

\begin{figure}
\includegraphics[width=\linewidth]{./images/fig_manipulations.png}
\caption{Surgical manipulations. Row A: wildtype. Row B: 90 degree
rotation. Row C: 180 degree rotation. Row D: Graft swap.}
\label{f:surg1}
\end{figure}

>>>>

% Allow includegraphics and nice wrapped figures
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage[outercaption]{sidecap}
\usepackage{subfigure}

@@

At each timestep in a simulation, the position, $\mathbf{x}_{b,t}$, at time
$t$ of each branch $b$ is updated according to
%
\begin{equation}
\mathbf{x}_{b,t+1} = \mathbf{x}_{b,t} + \mathbf{M}_{b}
\end{equation}
%
where $\mathbf{M}_{b}$, the movement vector, is the sum of a chemoaffinity
effect, $\mathbf{G}_b$, axon-axon competition, $\mathbf{C}_b$, an axon-axon
receptor-receptor interaction, $\mathbf{I}_b$, and axon-axon receptor-ligand
interaction, $\mathbf{J}_b$, and a `border effect',
$\mathbf{B}_b$, which acts to retain branches within the tectal region:
%
\begin{equation}
\mathbf{M}_{b} = m_1 \mathbf{G}_b + m_2 \mathbf{C}_b + m_3 \mathbf{I}_b +
m_4 \mathbf{J}_b + m_5 \mathbf{B}_b
\end{equation}
%
evaluated at time $t$. $m_1$--$m_5$ are scalar parameters.

@@

@@

\subsection*{Space based (or self-other) competition}

Competition, in the context of this model, is a repulsive effect which acts
between any two branches, whether they originate from different retinal
ganglion cells or from the same cell. It assumes that there is some kind of
self-other signal (\textbf{be sure to discuss possible origins}) which acts to
discourage two cells from approaching each other.
%
\begin{equation}
\mathbf{C}_b = \frac{1}{|B_{b,c}|} \sum_k \hat{\mathbf{x}}_{kb} W
\end{equation}
where the distance based weight $W$ is given by
\begin{equation}
W(d_{kb}) = \begin{cases}
      1 - \frac{d_{kb}}{2r}   & d_{kb} \leq 2r_c \\
     0 & d_{kb} > 2r_c
     \end{cases}
\end{equation}
for two growth cones of interaction radius $r_c$ separated by a distance
$d_{kb}$. $B_{b,c}$ is the set of growth cones within a distance $2r_c$ of
branch $b$.

\subsection*{Receptor-receptor axon-axon interactions}

We model interactions between axon growth cones due to signalling between
receptors of the same type. For these interactions, repulsion causes a
movement of branch $b$ along a unit vector, $\hat{\mathbf{x}}_{kb}$, from
branch $k$ to branch $b$ causing them to move further apart; attraction causes
the opposite movement. The interaction, $\mathbf{I}_b$, acting on branch $b$,
is given by

\textbf{FIXME} In the code, I'm mostly using the Reber-like ratio based
interaction, as described in Simpson and Goodhill.

%
\begin{equation}
% No \frac{1}{|B_b|} in this one. Really?
\mathbf{I}_b = \sum_k \hat{\mathbf{x}}_{kb}\,Q(d_{kb})
\end{equation}
%
where $Q_I(d_{kb})$ is the signalling strength between two growth cones of
receptor-receptor interaction radius $r_i$ a distance $d_{kb}$ from one
another, given by
%
\begin{equation}
Q_I(d_{kb}) = \begin{cases}
     \sum_i^N E_i\,r_{i,b}\,r_{i,k}    & d_{kb} \leq 2r_i \\
     0 & d_{kb} > 2r_i
     \end{cases}
\end{equation}
%
The sign of $Q_I$ is dependent on the values of $E_i$, which determines
whether the receptor-recetpro interaction, $\mathbf{I}_b$, is repulsive or
attractive. $r_{i,k}$ is the expression of receptor $i$ on branch $k$.

\subsection*{Receptor-ligand axon-axon interactions}

%
\begin{equation}
% No \frac{1}{|B_b|} in this one. Really?
\mathbf{J}_b = \sum_k \hat{\mathbf{x}}_{kb}\,Q_J(d_{kb})
\end{equation}
%
where $Q_J(d_{kb})$ is the signalling strength between two growth cones of
receptor-ligand interaction radius $r_j$ a distance $d_{kb}$ from one another,
given by
%
\begin{equation}
Q_J(d_{kb}) = \begin{cases}
     \sum_i^N F_i\,r_{i,b}\,l_{i,k}    & d_{kb} \leq 2r_j \\
     0 & d_{kb} > 2r_j
     \end{cases}
\end{equation}
%
The sign of $Q_J$ (which is dependent on the values of $F_i$) determines whether
the interaction, $\mathbf{I}_b$, is repulsive or attractive. $l_{i,k}$ is the
expression of ligand $i$ on branch $k$.

@@

\begin{figure}
\includegraphics[width=\linewidth]{./images/j4_eE_GJ_wt.png}
\caption{Gradient following model with axon-axon interactions.}
\label{f:wt}
\end{figure}

@@

This result shows that a gradient-following model based on receptor-ligand
interactions (both for axon-to-tectum and axon-to-axon interactions) is
able to generate the basic mapping observed in the retinotectal projection.

@@

%\includegraphics[width=\linewidth]{./images/fig_manipulations.png}

\begin{figure}
\begin{subfigure}{\linewidth}
\includegraphics[width=\linewidth]{./images/j4_eE_GC_wt_rotcmp.png}
\end{subfigure}
\begin{subfigure}{\linewidth}
\includegraphics[width=\linewidth]{./images/j4_eE_GC_tecrot90.png}
\end{subfigure}
\begin{subfigure}{\linewidth}
\includegraphics[width=\linewidth]{./images/j4_eE_GC_tecrot180.png}
\end{subfigure}
\begin{subfigure}{\linewidth}
\includegraphics[width=\linewidth]{./images/j4_eE_GC_tecswap.png}
\end{subfigure}
\caption{Surgical manipulations. Row A: wildtype. Row B: 90 degree
rotation. Row C: 180 degree rotation. Row D: Graft swap.}
\label{f:surg1}
\end{figure}
